---
title: "Fantasy Baseball WAR Project"
toc: false
about: 
  template: jolla
format:
  html:
    theme: lux
    css: styles.css
editor_options: 
  chunk_output_type: console
---


```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(baseballr)
library(mlbplotR)
library(dplyr)
library(tidytext)
library(tidyr)
library(ggplot2)
library(tidyverse)
```

In this personal project, I wrote code in R to calculate the value of player, given a fantasy baseball league's scoring system. WAR (Wins Above Replacement) is a statistic that Major League Baseball teams use to calculate the value a player adds over a replacement-level player at the same position. I applied the same concept to fantasy baseball in this project, where I identified:

1. Replacement-level players at every position: I found the level of replacement by looking at which players were on a team, and which players were listed as 'free agents.' Players listed as free agents may be added by any team, which aligns with the definition of replacement-level.

2. Points needed for a win: To win a game in fantasy baseball, a team must score more points than their opponent. To find the amount of points needed to win a matchup, I compiled a list of every matchup's winners and averaged their point totals. In theory, this stat describes 'what it takes' to win a game, on average. This value ended up being 447.29 points per game

Finally, to calculate the WAR, I started by determining how many wins a player creates, which I did by dividing their total points by the team points needed for a win. To find how many wins their replacement generated, I used the same process. The difference between the two values describes the "Wins Above Replacement" (or WAR) for a player. The equation is shown below:

WAR = (pts / 447.29) - (catcher_rep / 447.29)

The following is a sample of the top 50 players in the league during the 2024 MLB season, followed by a sample of the code I used to calculate WAR for both hitters and pitchers.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
Season <- 2024
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
catcher_hit <- mlb_stats(stat_type = 'season',
  position = "C",
  player_pool = 'all',
  stat_group = 'hitting',
  season = Season, 
  limit = 50) |>
  mutate(G = games_played,
         R = runs, 
         HR = home_runs,
         K = strike_outs,
         BB = base_on_balls,
         IBB = intentional_walks,
         H = hits,
         HBP = hit_by_pitch,
         AB = at_bats,
         SB = stolen_bases,
         GIDP = ground_into_double_play,
         PA = plate_appearances,
         TB = total_bases,
         RBI = rbi,
         BABIP = babip,
         name = player_full_name,
         Pos = "C") |>
  mutate(pts = (R*1 + RBI*1 + H*1 + TB*1 + BB*2 + IBB*1 - K*1 - GIDP*1 + SB*2)) |>
  select(name, pts, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP)

select_catchers <- catcher_hit |> pull(name)
         
catcher_field <- mlb_stats(stat_type = 'season',
  position = "C",
  player_pool = 'all',
  stat_group = 'fielding',
  season = Season, 
  limit = 750) |> 
  mutate(name = player_full_name,
         E = errors) |>
  group_by(name) |>
  summarize(E = sum(E, na.rm = TRUE)) |>
  select(name, E)

catcher_OFAST <- mlb_stats(stat_type = 'season',
  position = "C",
  player_pool = 'all',
  stat_group = 'fielding',
  season = Season, 
  limit = 750) |>
  mutate(name = player_full_name,
         OFAST = assists) |>
  filter(position_name == "Outfielder") |>
  select(name, OFAST) |>
  group_by(name) |>                            
  summarize(OFAST = sum(OFAST, na.rm = TRUE)) 

catchers <- catcher_hit |>
  left_join(catcher_field, by = "name") |>
  left_join(catcher_OFAST, by = "name") |>
  replace_na(list(OFAST = 0)) |>
  mutate(pts = (pts - E + OFAST*3), 
         ppg = pts/G) |>
  select(name, pts, ppg, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP, E, OFAST) |>
  arrange(desc(pts))
    
    
catcher_rep <- catchers |>                                   #Just Catchers in the year 2025
  ungroup() |>
  arrange(desc(pts), desc(ppg)) |>
  distinct() |>
  slice(15) |>                                                #replacement level
  pull(pts)
  
catchers <- catchers |>
  ungroup() |>
  arrange(desc(pts), desc(ppg)) |>
  mutate(WAR = round(((pts / (447.29 * 19)) - (catcher_rep / (447.29 * 19))) * 19, 3)
  ) |>
  select(name, WAR, pts, ppg, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP, E, OFAST) |>
  distinct()

rm(catcher_field, catcher_hit, catcher_OFAST, select_catchers)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
first_basemen_hit <- mlb_stats(stat_type = 'season',
  position = "1B",
  player_pool = 'all',
  stat_group = 'hitting',
  season = Season, 
  limit = 75) |>
  mutate(G = games_played,
         R = runs, 
         HR = home_runs,
         K = strike_outs,
         BB = base_on_balls,
         IBB = intentional_walks,
         H = hits,
         HBP = hit_by_pitch,
         AB = at_bats,
         SB = stolen_bases,
         GIDP = ground_into_double_play,
         PA = plate_appearances,
         TB = total_bases,
         RBI = rbi,
         BABIP = babip,
         name = player_full_name,
         Pos = "1B") |>
  mutate(pts = (R*1 + RBI*1 + H*1 + TB*1 + BB*2 + IBB*1 - K*1 - GIDP*1 + SB*2)) |>
  select(name, pts, G, Pos, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP)

select_first_basemen <- first_basemen_hit |> pull(name)
         
first_basemen_field <- mlb_stats(stat_type = 'season',
  position = "1B",
  player_pool = 'all',
  stat_group = 'fielding',
  season = Season, 
  limit = 750) |> 
  filter(position_abbreviation == "1B") |>
  mutate(name = player_full_name,
         E = errors) |>
  group_by(name) |>
  summarize(E = sum(E, na.rm = TRUE)) |>
  select(name, E)

first_basemen_OFAST <- mlb_stats(stat_type = 'season',
  position = "1B",
  player_pool = 'all',
  stat_group = 'fielding',
  season = Season, 
  limit = 750) |>
  mutate(name = player_full_name,
         OFAST = assists) |>
  filter(position_name == "Outfielder") |>
  select(name, OFAST) |>
  group_by(name) |>                            
  summarize(OFAST = sum(OFAST, na.rm = TRUE)) 

first_basemen <- first_basemen_hit |>
  left_join(first_basemen_field, by = "name") |>
  left_join(first_basemen_OFAST, by = "name") |>
  replace_na(list(OFAST = 0)) |>
  mutate(pts = (pts - E + OFAST*3), 
         ppg = pts/G) |>
  select(name, pts, ppg, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP, E, OFAST) |>
  arrange(desc(pts))
    
    
first_basemen_rep <- first_basemen |>                               #Just 1B in the year 2025
  ungroup() |>
  arrange(desc(pts), desc(ppg)) |>
  distinct() |>
  slice(22) |>                                                      #replacement level
  pull(pts)
  
first_basemen <- first_basemen |>
  ungroup() |>
  arrange(desc(pts), desc(ppg)) |>
  mutate(WAR = round(((pts / (447.29 * 19)) - (first_basemen_rep / (447.29 * 19))) * 19, 3)
  ) |>           #adjust
  select(name, WAR, pts, ppg, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP, E, OFAST) |>
  distinct()

rm(first_basemen_field, first_basemen_hit, first_basemen_OFAST, select_first_basemen)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
second_basemen_hit <- mlb_stats(stat_type = 'season',
  position = "2B",
  player_pool = 'all',
  stat_group = 'hitting',
  season = Season, 
  limit = 75) |>
  mutate(G = games_played,
         R = runs, 
         HR = home_runs,
         K = strike_outs,
         BB = base_on_balls,
         IBB = intentional_walks,
         H = hits,
         HBP = hit_by_pitch,
         AB = at_bats,
         SB = stolen_bases,
         GIDP = ground_into_double_play,
         PA = plate_appearances,
         TB = total_bases,
         RBI = rbi,
         BABIP = babip,
         name = player_full_name,
         Pos = "2B") |>
  mutate(pts = (R*1 + RBI*1 + H*1 + TB*1 + BB*2 + IBB*1 - K*1 - GIDP*1 + SB*2)) |>
  select(name, pts, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP)

select_second_basemen <- second_basemen_hit |> pull(name)
         
second_basemen_field <- mlb_stats(stat_type = 'season',
  position = "2B",
  player_pool = 'all',
  stat_group = 'fielding',
  season = Season, 
  limit = 700) |> 
  mutate(name = player_full_name,
         E = errors) |>
  group_by(name) |>
  summarize(E = sum(E, na.rm = TRUE)) |>
  select(name, E)

second_basemen_OFAST <- mlb_stats(stat_type = 'season',
  position = "2B",
  player_pool = 'all',
  stat_group = 'fielding',
  season = Season, 
  limit = 750) |>
  mutate(name = player_full_name,
         OFAST = assists) |>
  filter(position_name == "Outfielder") |>
  select(name, OFAST) |>
  group_by(name) |>                            
  summarize(OFAST = sum(OFAST, na.rm = TRUE)) 

second_basemen <- second_basemen_hit |>
  left_join(second_basemen_field, by = "name") |>
  left_join(second_basemen_OFAST, by = "name") |>
  replace_na(list(OFAST = 0)) |>
  mutate(pts = (pts - E + OFAST*3), 
         ppg = pts/G) |>
  select(name, pts, ppg, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP, E, OFAST) |>
  arrange(desc(pts))
    
    
second_basemen_rep <- second_basemen |>                               #Just 1B in the year 2025
  ungroup() |>
  arrange(desc(pts), desc(ppg)) |>
  distinct() |>
  slice(14) |>                                                      #replacement level
  pull(pts)
  
second_basemen <- second_basemen |>
  ungroup() |>
  arrange(desc(pts), desc(ppg)) |>
  mutate(WAR = round(((pts / (447.29 * 19)) - (second_basemen_rep / (447.29 * 19))) * 19, 3)
  ) |>       #adjust
  select(name, WAR, pts, ppg, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP, E, OFAST) |>
  distinct()

rm(second_basemen_field, second_basemen_hit, second_basemen_OFAST, select_second_basemen)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
shortstop_hit <- mlb_stats(stat_type = 'season',
  position = "SS",
  player_pool = 'all',
  stat_group = 'hitting',
  season = Season, 
  limit = 75) |>
  mutate(G = games_played,
         R = runs, 
         HR = home_runs,
         K = strike_outs,
         BB = base_on_balls,
         IBB = intentional_walks,
         H = hits,
         HBP = hit_by_pitch,
         AB = at_bats,
         SB = stolen_bases,
         GIDP = ground_into_double_play,
         PA = plate_appearances,
         TB = total_bases,
         RBI = rbi,
         BABIP = babip,
         name = player_full_name,
         Pos = "SS") |>
  mutate(pts = (R*1 + RBI*1 + H*1 + TB*1 + BB*2 + IBB*1 - K*1 - GIDP*1 + SB*2)) |>
  select(name, pts, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP)

select_shortstop <- shortstop_hit |> pull(name)
         
shortstop_field <- mlb_stats(stat_type = 'season',
  position = "SS",
  player_pool = 'all',
  stat_group = 'fielding',
  season = Season, 
  limit = 750) |> 
  mutate(name = player_full_name,
         E = errors) |>
  group_by(name) |>
  summarize(E = sum(E, na.rm = TRUE)) |>
  select(name, E)

shortstop_OFAST <- mlb_stats(stat_type = 'season',
  position = "SS",
  player_pool = 'all',
  stat_group = 'fielding',
  season = Season, 
  limit = 750) |>
  mutate(name = player_full_name,
         OFAST = assists) |>
  filter(position_name == "Outfielder") |>
  select(name, OFAST) |>
  group_by(name) |>                            
  summarize(OFAST = sum(OFAST, na.rm = TRUE)) 

shortstop <- shortstop_hit |>
  left_join(shortstop_field, by = "name") |>
  left_join(shortstop_OFAST, by = "name") |>
  replace_na(list(OFAST = 0)) |>
  mutate(pts = (pts - E + OFAST*3), 
         ppg = pts/G) |>
  select(name, pts, ppg, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP, E, OFAST) |>
  arrange(desc(pts))
    
    
shortstop_rep <- shortstop |>                               #Just 1B in the year 2025
  ungroup() |>
  arrange(desc(pts), desc(ppg)) |>
  distinct() |>
  slice(21) |>                                                      #replacement level
  pull(pts)
  
shortstop <- shortstop |>
  ungroup() |>
  arrange(desc(pts), desc(ppg)) |>
  mutate(WAR = round(((pts / (447.29 * 19)) - (shortstop_rep / (447.29 * 19))) * 19, 3)
  ) |>         #adjust
  select(name, WAR, pts, ppg, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP, E, OFAST) |>
  distinct()

rm(shortstop_field, shortstop_hit, shortstop_OFAST, select_shortstop)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
third_basemen_hit <- mlb_stats(stat_type = 'season',
  position = "3B",
  player_pool = 'all',
  stat_group = 'hitting',
  season = Season, 
  limit = 75) |>
  mutate(G = games_played,
         R = runs, 
         HR = home_runs,
         K = strike_outs,
         BB = base_on_balls,
         IBB = intentional_walks,
         H = hits,
         HBP = hit_by_pitch,
         AB = at_bats,
         SB = stolen_bases,
         GIDP = ground_into_double_play,
         PA = plate_appearances,
         TB = total_bases,
         RBI = rbi,
         BABIP = babip,
         name = player_full_name,
         Pos = "3B") |>
  mutate(pts = (R*1 + RBI*1 + H*1 + TB*1 + BB*2 + IBB*1 - K*1 - GIDP*1 + SB*2)) |>
  select(name, pts, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP)

select_third_basemen <- third_basemen_hit |> pull(name)
         
third_basemen_field <- mlb_stats(stat_type = 'season',
  position = "3B",
  player_pool = 'all',
  stat_group = 'fielding',
  season = Season, 
  limit = 750) |> 
  mutate(name = player_full_name,
         E = errors) |>
  group_by(name) |>
  summarize(E = sum(E, na.rm = TRUE)) |>
  select(name, E)

third_basemen_OFAST <- mlb_stats(stat_type = 'season',
  position = "3B",
  player_pool = 'all',
  stat_group = 'fielding',
  season = Season, 
  limit = 750) |>
  mutate(name = player_full_name,
         OFAST = assists) |>
  filter(position_name == "Outfielder") |>
  select(name, OFAST) |>
  group_by(name) |>                            
  summarize(OFAST = sum(OFAST, na.rm = TRUE)) 

third_basemen <- third_basemen_hit |>
  left_join(third_basemen_field, by = "name") |>
  left_join(third_basemen_OFAST, by = "name") |>
  replace_na(list(OFAST = 0)) |>
  mutate(pts = (pts - E + OFAST*3), 
         ppg = pts/G) |>
  select(name, pts, ppg, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP, E, OFAST) |>
  arrange(desc(pts))
    
    
third_basemen_rep <- third_basemen |>                               #Just 1B in the year 2025
  ungroup() |>
  arrange(desc(pts), desc(ppg)) |>
  distinct() |>
  slice(15) |>                                                      #replacement level
  pull(pts)
  
third_basemen <- third_basemen |>
  ungroup() |>
  arrange(desc(pts), desc(ppg)) |>
  mutate(WAR = round(((pts / (447.29 * 19)) - (third_basemen_rep / (447.29 * 19))) * 19, 3)
  ) |>       #adjust
  select(name, WAR, pts, ppg, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP, E, OFAST) |>
  distinct()

rm(third_basemen_field, third_basemen_hit, third_basemen_OFAST, select_third_basemen)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
outfielder_hit <- mlb_stats(stat_type = 'season',
  position = "OF",
  player_pool = 'all',
  stat_group = 'hitting',
  season = Season, 
  limit = 150) |>
  mutate(G = games_played,
         R = runs, 
         HR = home_runs,
         K = strike_outs,
         BB = base_on_balls,
         IBB = intentional_walks,
         H = hits,
         HBP = hit_by_pitch,
         AB = at_bats,
         SB = stolen_bases,
         GIDP = ground_into_double_play,
         PA = plate_appearances,
         TB = total_bases,
         RBI = rbi,
         BABIP = babip,
         name = player_full_name, 
         Pos = "OF") |>
  mutate(pts = (R*1 + RBI*1 + H*1 + TB*1 + BB*2 + IBB*1 - K*1 - GIDP*1 + SB*2)) |>
  select(name, pts, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP)

select_outfielder <- outfielder_hit |> pull(name)
         
outfielder_field <- mlb_stats(stat_type = 'season',
  position = "OF",
  player_pool = 'all',
  stat_group = 'fielding',
  season = Season, 
  limit = 750) |> 
  mutate(name = player_full_name,
         E = errors) |>
  group_by(name) |>
  summarize(E = sum(E, na.rm = TRUE))
  
outfielder_assist <- mlb_stats(stat_type = 'season',
  position = "OF",
  player_pool = 'all',
  stat_group = 'fielding',
  season = Season, 
  limit = 750) |>
  mutate(name = player_full_name,
         OFAST = assists) |>
  filter(position_name == "Outfielder") |>
  select(name, OFAST) |>
  group_by(name) |>                           
  summarize(OFAST = sum(OFAST, na.rm = TRUE)) 

outfielder <- outfielder_hit |>
  left_join(outfielder_field, by = "name") |>
  left_join(outfielder_assist, by = "name") |>
  replace_na(list(OFAST = 0)) |>
  mutate(pts = (pts - E + OFAST*3), 
         ppg = pts/G) |>
  select(name, pts, ppg, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP, E, OFAST) |>
  arrange(desc(pts))
    
    
outfielder_rep <- outfielder |>                               #Just 1B in the year 2025
  ungroup() |>
  arrange(desc(pts), desc(ppg)) |>
  distinct() |>
  slice(55) |>                                                      #replacement level
  pull(pts)
  
outfielder <- outfielder |>
  ungroup() |>
  arrange(desc(pts), desc(ppg)) |>
  mutate(WAR = round(((pts / (447.29 * 19)) - (outfielder_rep / (447.29 * 19))) * 19, 3)
  ) |>          #adjust
  select(name, WAR, pts, ppg, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP, E, OFAST) |>
  distinct()

rm(outfielder_field, outfielder_hit, outfielder_assist, select_outfielder)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
utility_hit <- mlb_stats(stat_type = 'season',
  position = "DH",
  player_pool = 'all',
  stat_group = 'hitting',
  season = Season, 
  limit = 50) |>
  mutate(G = games_played,
         R = runs, 
         HR = home_runs,
         K = strike_outs,
         BB = base_on_balls,
         IBB = intentional_walks,
         H = hits,
         HBP = hit_by_pitch,
         AB = at_bats,
         SB = stolen_bases,
         GIDP = ground_into_double_play,
         PA = plate_appearances,
         TB = total_bases,
         RBI = rbi,
         BABIP = babip,
         name = player_full_name, 
         Pos = "UTIL") |>
  mutate(pts = (R*1 + RBI*1 + H*1 + TB*1 + BB*2 + IBB*1 - K*1 - GIDP*1 + SB*2)) |>
  select(name, pts, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP)

select_utility <- utility_hit |> pull(name)

utility_OFAST <- mlb_stats(stat_type = 'season',
  position = "DH",
  player_pool = 'all',
  stat_group = 'fielding',
  season = Season, 
  limit = 750) |>
  mutate(name = player_full_name,
         OFAST = assists) |>
  filter(position_name == "Outfielder") |>
  select(name, OFAST) |>
  group_by(name) |>
  summarize(OFAST = max(OFAST))

utility <- utility_hit |>
  left_join(utility_OFAST, by = "name") |>
  replace_na(list(OFAST = 0)) |>
  mutate(pts = (pts + OFAST*3), 
         ppg = pts/G) |>
  select(name, pts, ppg, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP, OFAST) |>
  arrange(desc(pts))
    
    
utility_rep <- max(c(`catcher_rep`, `first_basemen_rep`, `second_basemen_rep`, shortstop_rep, third_basemen_rep, outfielder_rep))                        #top positional weight
  
utility <- utility |>
  ungroup() |>
  arrange(desc(pts), desc(ppg)) |>
  mutate(WAR = round(((pts / (447.29 * 19)) - (utility_rep / (447.29 * 19))) * 19, 3)
  ) |>           #adjust
  select(name, WAR, pts, ppg, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP, OFAST) |>
  distinct()

rm(utility_hit, utility_OFAST, select_utility)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
WAR_Hit <- catchers |>
  full_join(first_basemen) |>
  full_join(second_basemen) |>
  full_join(shortstop) |>
  full_join(third_basemen) |>
  full_join(outfielder) |>
  full_join(utility) |>
  replace_na(list(E = 0)) |>
  arrange(desc(WAR))
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
convert_innings <- function(ip) {
  whole <- floor(ip)
  decimal <- ip - whole
  if (decimal >= 0.3) return(NA_real_)  # to avoid bad conversions
  return(whole + (decimal * 10) / 3)}

starting_pitchers <- mlb_stats(stat_type = 'season',
  position = "P",
  player_pool = 'all',
  stat_group = 'pitching',
  season = Season, 
  limit = 750) |>
  mutate(G = games_played,
         GS = games_started,
         GO = ground_outs,
         FO = air_outs,
         HR = home_runs,
         K = strike_outs,
         BB = base_on_balls,
         IBB = intentional_walks,
         H = hits,
         HBP = hit_by_pitch,
         AB = at_bats,
         GIDP = ground_into_double_play,
         Pitches = number_of_pitches,
         IP_raw = as.numeric(innings_pitched),
         SV = saves,
         HD = holds,
         BS = blown_saves,
         ER = earned_runs,
         `Strike %` = strike_percentage,
         B = balks,
         PKO = pickoffs, 
         name = player_full_name,
         Pos = "SP") |>
  mutate(IP = sapply(IP_raw, convert_innings)) |>
  mutate(pts = (IP*4.5 - H*1 - ER*3 - BB*1 + K*1 - B*1 + PKO*1 + SV*6 + HD*5 - BS*3),
         ppg = pts/G) |>
  select(name, pts, ppg, Pos, G, GS, IP, K, BB, H, ER, SV, HD, BS, B, PKO, `Strike %`, Pitches, AB, GIDP, HBP, IBB, GO, FO) |>
  filter((Pitches/G) >= 30) |>
  arrange(desc(pts))

select_pitchers <- starting_pitchers |> pull(name)
    
starting_pitcher_rep <- starting_pitchers |>                          #Just SP in the year 2025
  ungroup() |>
  arrange(desc(pts), desc(ppg)) |>
  distinct() |>
  slice(92) |>                                                      #replacement level
  pull(pts)
  
starting_pitchers <- starting_pitchers |>
  ungroup() |>
  arrange(desc(pts), desc(ppg)) |>
 mutate(WAR = round(((pts / (447.29 * 19)) - (starting_pitcher_rep / (447.29 * 19))) * 19, 3)
  ) |>         #adjust
  select(name, WAR, pts, ppg, Pos, G, GS, IP, K, BB, H, ER, SV, HD, BS, B, PKO, `Strike %`, Pitches, AB, GIDP, HBP, IBB, GO, FO) |>
  distinct()
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
convert_innings <- function(ip) {
  whole <- floor(ip)
  decimal <- ip - whole
  if (decimal >= 0.3) return(NA_real_)  # to avoid bad conversions
  return(whole + (decimal * 10) / 3)}

relief_pitchers <- mlb_stats(stat_type = 'season',
  position = "P",
  player_pool = 'all',
  stat_group = 'pitching',
  season = Season, 
  limit = 600) |>
  mutate(G = games_played,
         GS = games_started,
         GO = ground_outs,
         FO = air_outs,
         HR = home_runs,
         K = strike_outs,
         BB = base_on_balls,
         IBB = intentional_walks,
         H = hits,
         HBP = hit_by_pitch,
         AB = at_bats,
         GIDP = ground_into_double_play,
         Pitches = number_of_pitches,
         IP_raw = as.numeric(innings_pitched),
         SV = saves,
         HD = holds,
         BS = blown_saves,
         ER = earned_runs,
         `Strike %` = strike_percentage,
         B = balks,
         PKO = pickoffs, 
         name = player_full_name,
         Pos = "RP") |>
  mutate(IP = sapply(IP_raw, convert_innings)) |>
  mutate(pts = (IP*4.5 - H*1 - ER*3 - BB*1 + K*1 - B*1 + PKO*1 + SV*6 + HD*5 - BS*3),
         ppg = pts/G) |>
  select(name, pts, ppg, Pos, G, GS, IP, K, BB, H, ER, SV, HD, BS, B, PKO, `Strike %`, Pitches, AB, GIDP, HBP, IBB, GO, FO) |>
  filter((Pitches/G) <= 30) |>
  arrange(desc(pts)) |>
  slice_max(order_by = pts, n = 150)

select_relief_pitchers <- relief_pitchers |> pull(name)
    
relief_pitcher_rep <- relief_pitchers |>                          #Just SP in the year 2025
  ungroup() |>
  arrange(desc(pts), desc(ppg)) |>
  distinct() |>
  slice(26) |>                                                      #replacement level
  pull(pts)
  
relief_pitchers <- relief_pitchers |>
  ungroup() |>
  arrange(desc(pts), desc(ppg)) |>
  mutate(WAR = round(((pts / (447.29 * 19)) - (relief_pitcher_rep / (447.29 * 19))) * 19, 3)
  ) |>
  select(name, WAR, pts, ppg, Pos, G, GS, IP, K, BB, H, ER, SV, HD, BS, B, PKO, `Strike %`, Pitches, AB, GIDP, HBP, IBB, GO, FO) |>
  distinct()
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
WAR_Pitch <- starting_pitchers |>
  full_join(relief_pitchers) |>
  arrange(desc(WAR))
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
WAR_All <- WAR_Pitch |>
  full_join(WAR_Hit) |>
  select(name, WAR, pts, ppg, Pos, G) |>
  arrange(desc(WAR))
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
statcast_hit_exitvelo <-scrape_savant_leaderboards(
  leaderboard = "exit_velocity_barrels",
  year = Season,
  abs = 20,
  min_pa = "q",
  min_pitches = 50,
  min_throws = 5,
  min_field = "q",
  min_run = 0,
  player_type = "batter",
  fielding_type = "player",
  oaa_position = "",
  oaa_roles = "",
  team = "",
  arsenal_type = "n_",
  run_type = "raw",
  min2b = 5,
  min3b = 0,
  position = "",
  bats = "",
  hand = ""
) |>
  separate(`last_name, first_name`, into = c("last", "first"), sep = ",\\s*") |> 
  mutate(name = paste(first, last)) |>
  mutate(launch_angle = avg_hit_angle,
         sweetspot = anglesweetspotpercent) |>
  select(name, launch_angle, sweetspot, avg_hit_speed, ev95percent, brl_pa)

statcast_hit_exp <-scrape_savant_leaderboards(
  leaderboard = "expected_statistics",
  year = Season,
  abs = 20,
  min_pa = "q",
  min_pitches = 50,
  min_throws = 5,
  min_field = "q",
  min_run = 0,
  player_type = "batter",
  fielding_type = "player",
  oaa_position = "",
  oaa_roles = "",
  team = "",
  arsenal_type = "n_",
  run_type = "raw",
  min2b = 5,
  min3b = 0,
  position = "",
  bats = "",
  hand = ""
) |>
  separate(`last_name, first_name`, into = c("last", "first"), sep = ",\\s*") |> 
  mutate(name = paste(first, last)) |>
  mutate(ba_diff = est_ba_minus_ba_diff,
         slg_diff = est_slg_minus_slg_diff,
         woba_diff = est_woba_minus_woba_diff) |>
  select(name, ba_diff, slg_diff, woba_diff)

WAR_Hit_Statcast <- WAR_Hit |>
  left_join(statcast_hit_exp) |>
  left_join(statcast_hit_exitvelo)

#write_csv(WAR_Hit_Statcast, "2025_hit.csv")
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
Breakout_Hit <- WAR_Hit_Statcast |>
  mutate(BABIP = as.numeric(BABIP)) |>
  filter(ppg >= 2.6,
         woba_diff <= 0,
         ba_diff <= 0,
         slg_diff <= 0) |>
mutate(improve_hit = ((woba_diff + ba_diff + slg_diff - 1)*-1)*(.3/BABIP)) |>
  arrange(desc(improve_hit)) |>
  mutate(newppg = ppg*(sqrt(improve_hit)))
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
Crashout_Hit <- WAR_Hit_Statcast |>
  mutate(BABIP = as.numeric(BABIP)) |>
  filter(ppg >= 2.6,
         woba_diff >= 0,
         ba_diff >= 0,
         slg_diff >= 0) |>
mutate(crash_hit = (woba_diff + ba_diff + slg_diff + 1)*(BABIP/.3)) |>
  arrange(desc(crash_hit)) |>
  mutate(newppg = ppg/(sqrt(crash_hit)))
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
statcast_pitch_exitvelo <-scrape_savant_leaderboards(
  leaderboard = "exit_velocity_barrels",
  year = Season,
  abs = 20,
  min_pa = "q",
  min_pitches = 50,
  min_throws = 5,
  min_field = "q",
  min_run = 0,
  player_type = "pitcher",
  fielding_type = "player",
  oaa_position = "",
  oaa_roles = "",
  team = "",
  arsenal_type = "n_",
  run_type = "raw",
  min2b = 5,
  min3b = 0,
  position = "",
  bats = "",
  hand = ""
) |>
  separate(`last_name, first_name`, into = c("last", "first"), sep = ",\\s*") |> 
  mutate(name = paste(first, last)) |>
  mutate(launch_angle = avg_hit_angle,
         sweetspot = anglesweetspotpercent) |>
  select(name, launch_angle, sweetspot, avg_hit_speed, ev95percent, brl_pa)

statcast_pitch_exp <-scrape_savant_leaderboards(
  leaderboard = "expected_statistics",
  year = Season,
  abs = 20,
  min_pa = "q",
  min_pitches = 50,
  min_throws = 5,
  min_field = "q",
  min_run = 0,
  player_type = "pitcher",
  fielding_type = "player",
  oaa_position = "",
  oaa_roles = "",
  team = "",
  arsenal_type = "n_",
  run_type = "raw",
  min2b = 5,
  min3b = 0,
  position = "",
  bats = "",
  hand = ""
) |>
  separate(`last_name, first_name`, into = c("last", "first"), sep = ",\\s*") |> 
  mutate(name = paste(first, last)) |>
  mutate(ba_diff = est_ba_minus_ba_diff,
         slg_diff = est_slg_minus_slg_diff,
         woba_diff = est_woba_minus_woba_diff,
         era_diff = era_minus_xera_diff) |>
  select(name, ba_diff, slg_diff, woba_diff, era_diff)

WAR_Pitch_Statcast <- WAR_Pitch |>
  left_join(statcast_pitch_exp) |>
  left_join(statcast_pitch_exitvelo)

#write_csv(WAR_Pitch_Statcast, "2025_pitch.csv")
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
Breakout_Pitch_SP <- WAR_Pitch_Statcast |>
  filter(Pos == "SP") |>
  filter(ppg >= 8,
         woba_diff >= 0,
         ba_diff >= 0,
         slg_diff >= 0, 
         era_diff >= 0) |>
mutate(improve_SP = ((woba_diff + ba_diff + slg_diff + 1)*((era_diff/25)+1))^2) |>
  arrange(desc(improve_SP)) |>
  mutate(newppg = ppg*(sqrt(improve_SP)))
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
Crashout_Pitch_SP <- WAR_Pitch_Statcast |>
  filter(ppg >= 8,
         woba_diff <= 0,
         ba_diff <= 0,
         slg_diff <= 0,
         era_diff <= 0) |>
mutate(crash_SP = (((woba_diff + ba_diff + slg_diff - 1)*-1)*((era_diff/-25)+1))^2) |>
  arrange(desc(crash_SP)) |>
  mutate(newppg = ppg/(sqrt(crash_SP)))
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
WAR_All |> print(n = 50)
```

# Code Sample for Hitters

```{r, message = FALSE, warning = FALSE}
catcher_hit <- mlb_stats(stat_type = 'season',          #pulls data using mlb_stats package
  position = "C",                                       #selects only Catchers
  player_pool = 'all',
  stat_group = 'hitting',                               #pulls only hitting statistics
  season = Season,                                      
  limit = 50) |>                                        #takes only the top 50 catchers
  mutate(G = games_played,                              #renames stats to abbreviations
         R = runs, 
         HR = home_runs,
         K = strike_outs,
         BB = base_on_balls,
         IBB = intentional_walks,
         H = hits,
         HBP = hit_by_pitch,
         AB = at_bats,
         SB = stolen_bases,
         GIDP = ground_into_double_play,
         PA = plate_appearances,
         TB = total_bases,
         RBI = rbi,
         BABIP = babip,
         name = player_full_name,
         Pos = "C") |>
  mutate(pts = (R*1 + RBI*1 + H*1 + TB*1 + BB*2 + IBB*1 - K*1 - GIDP*1 + SB*2)) |>   #calculates points for a player, given the league scoring system for each hitting statistic
  select(name, pts, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP)
         
catcher_field <- mlb_stats(stat_type = 'season',
  position = "C",
  player_pool = 'all',
  stat_group = 'fielding',                            #selects fielding statistics
  season = Season, 
  limit = 750) |> 
  mutate(name = player_full_name,                     #renames columns succinctly
         E = errors) |>
  group_by(name) |>
  summarize(E = sum(E, na.rm = TRUE)) |>
  select(name, E)                                     #keeps only player name and Errors

catchers <- catcher_hit |>                            #combines data frames
  left_join(catcher_field, by = "name") |>
  mutate(pts = (pts - E),                        #calculates total points with fielding errors
         ppg = pts/G) |>                         #calculates points per game
  select(name, pts, ppg, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP, E) |>
  arrange(desc(pts))                                  #arranges column by points
    
    
catcher_rep <- catchers |>                            #saves replacement-level catcher
  ungroup() |>
  arrange(desc(pts), desc(ppg)) |>
  distinct() |>
  slice(15) |>                                       #15th best catcher is 'replacement-level'
  pull(pts)
  
catchers <- catchers |>                              #edits the 'catchers' data frame
  ungroup() |>
  arrange(desc(pts), desc(ppg)) |>
  mutate(WAR = round((pts / 447.29) - (catcher_rep / 447.29)), 3) |>             #finds a player's WAR in a year, given that 447.29 points creates a win
  select(name, WAR, pts, ppg, Pos, G, PA, AB, H, TB, RBI, R, HR, BB, IBB, K, SB, GIDP, HBP, BABIP, E) |>
  distinct()

rm(catcher_field, catcher_hit)                #cleans environment in R
```

# Code Sample for Pitchers

```{r, message = FALSE, warning = FALSE}
convert_innings <- function(ip) {
  whole <- floor(ip)                    #finds whole number in IP statistic
  decimal <- ip - whole                 #finds the decimal in the IP statistic
  if (decimal >= 0.3) return(NA_real_)  #avoids any bad conversions
  return(whole + (decimal * 10) / 3)}   #recalculates decimal to be more statistically accurate

starting_pitchers <- mlb_stats(stat_type = 'season',
  position = "P",                        #finds only pitchers
  player_pool = 'all',
  stat_group = 'pitching',
  season = Season, 
  limit = 750) |>
  mutate(G = games_played,
         GS = games_started,
         GO = ground_outs,
         FO = air_outs,
         HR = home_runs,
         K = strike_outs,
         BB = base_on_balls,
         IBB = intentional_walks,
         H = hits,
         HBP = hit_by_pitch,
         AB = at_bats,
         GIDP = ground_into_double_play,
         Pitches = number_of_pitches,
         IP_raw = as.numeric(innings_pitched),
         SV = saves,
         HD = holds,
         BS = blown_saves,
         ER = earned_runs,
         `Strike %` = strike_percentage,
         B = balks,
         PKO = pickoffs, 
         name = player_full_name,
         Pos = "SP") |>
  mutate(IP = sapply(IP_raw, convert_innings)) |>     #uses function to convert IP
  mutate(pts = (IP*4.5 - H*1 - ER*3 - BB*1 + K*1 - B*1 + PKO*1 + SV*6 + HD*5 - BS*3),
         ppg = pts/G) |>              #calculates points using league's scoring system
  select(name, pts, ppg, Pos, G, GS, IP, K, BB, H, ER, SV, HD, BS, B, PKO, `Strike %`, Pitches, AB, GIDP, HBP, IBB, GO, FO) |>
  filter((Pitches/G) >= 30) |>        #only takes pitchers that tend to throw more than 30 pitchers per game (relief pitchers will throw under 30, starters will throw over 30)
  arrange(desc(pts))
    
starting_pitcher_rep <- starting_pitchers |>        #saves replacement-level starting pitcher
  ungroup() |>
  arrange(desc(pts), desc(ppg)) |>
  distinct() |>
  slice(92) |>                              #92nd best starting pitcher is 'replacement-level'
  pull(pts)
  
starting_pitchers <- starting_pitchers |>      #edits the starting_pitchers data frame
  ungroup() |>
  arrange(desc(pts), desc(ppg)) |>
 mutate(WAR = round((pts / 447.29) - (starting_pitcher_rep / 447.29)), 3) |>  #finds a player's WAR in a year, given that 447.29 points creates a win
  select(name, WAR, pts, ppg, Pos, G, GS, IP, K, BB, H, ER, SV, HD, BS, B, PKO, `Strike %`, Pitches, AB, GIDP, HBP, IBB, GO, FO) |>
  distinct()
```
```

