---
title: "2026 Senate Forecast"
toc: false
about:
  template: jolla
format:
  html:
    theme: lux
editor_options:
  chunk_output_type: console
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(rvest)
library(polite)
library(purrr)
library(httr2)
library(httr)
library(lubridate)
library(maps)
library(statebins)
library(sf)
library(leaflet)
library(htmltools)
```

This model forecasts the 2026 U.S. Senate Elections for 13 key races.

**D - 52, R - 48**

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#National Democratic Advantage
D_Adv <- 7

#Importance of National Environment
Nat_Env = .15

#Importance of State Environment
Sta_Env = .15
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#| include: FALSE
#State Cook Political Voting Index
robotstxt::paths_allowed("https://en.wikipedia.org/wiki/Cook_Partisan_Voting_Index")
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#| include: FALSE
#Scrape
CPVI <- read_html("https://en.wikipedia.org/wiki/Cook_Partisan_Voting_Index")
str(CPVI)

#Get Table
table <- html_nodes(CPVI, css = "table")
untidy_tables <- html_table(table, header = TRUE, fill = TRUE)
StatesCPVI <- untidy_tables[[2]]
CPVI <- StatesCPVI |>
  mutate(state = state.abb[match(StatesCPVI$State, state.name)]) |>
  mutate(
    pvi = case_when(
      grepl("EVEN", PVI, ignore.case = TRUE) ~ 0,
      grepl("^D", PVI) ~ as.numeric(sub("D\\+", "", PVI)),
      grepl("^R", PVI) ~ -as.numeric(sub("R\\+", "", PVI)))) |>
  mutate(adj_pvi = pvi + D_Adv) |>
  select(state, pvi, adj_pvi)

rm(StatesCPVI, untidy_tables, table)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
today <- today()

polls_unclean <- read.csv("~/Downloads/senate.csv")
polls <- polls_unclean |>
  select(state, end_date, sponsor_candidate_party, sample_size, population, stage, pct, answer, poll_id) |>
  mutate(date = mdy(end_date)) |>
  filter(pct >= 10)

AK_polls <- polls |>
  filter(state == "AK") |>
  filter(stage == "general") |>
  filter(answer %in% c("Peltola", "Sullivan (Dan)")) |>
    group_by(poll_id) |>
    filter(n_distinct(answer) > 1) |> 
    distinct(answer, .keep_all = TRUE) |> 
    slice_head(n = 2) |>
   ungroup() |>
pivot_wider(names_from = "answer", values_from = "pct") |>
  rename("Republican" = "Sullivan (Dan)") |>
  rename("Democrat" = "Peltola") |>
  mutate(days = as.numeric(today() - date),
         weight = factor(
            case_when(
              days >= 0  & days <= 14 ~ "4",
              days > 14  & days <= 31 ~ "3",
              days > 31  & days <= 90 ~ "2",
              days > 90 ~ "1"))) |>
  select(state, sample_size, population, Republican, Democrat, weight)

FL_polls <- polls |>
  filter(state == "FL") |>
  filter(stage == "general") |>
  filter(!answer %in% c("Vindman (Alexander)", "Don't know", "Jenkins", "Mujica"))  |>
  group_by(poll_id) |>
    filter(n_distinct(answer) > 1) |> 
    distinct(answer, .keep_all = TRUE) |> 
    slice_head(n = 2) |>
   ungroup() |>
pivot_wider(names_from = "answer", values_from = "pct") |>
  rename("Republican" = "Moody (Ashley)") |>
  rename("Democrat" = "Nixon (Angie)") |>
  mutate(days = as.numeric(today() - date),
         weight = factor(
            case_when(
              days >= 0  & days <= 14 ~ "4",
              days > 14  & days <= 31 ~ "3",
              days > 31  & days <= 90 ~ "2",
              days > 90 ~ "1"))) |>
  select(state, sample_size, population, Republican, Democrat, weight)

GA_polls <- polls |>
  filter(state == "GA") |>
  filter(stage == "general") |>
  filter(answer %in% c("Ossoff", "Collins (Mike)")) |>
  group_by(poll_id) |>
    filter(n_distinct(answer) > 1) |> 
    distinct(answer, .keep_all = TRUE) |> 
    slice_head(n = 2) |>
   ungroup() |>
pivot_wider(names_from = "answer", values_from = "pct") |>
  rename("Republican" = "Collins (Mike)") |>
  rename("Democrat" = "Ossoff") |>
  mutate(days = as.numeric(today() - date),
         weight = factor(
            case_when(
              days >= 0  & days <= 14 ~ "4",
              days > 14  & days <= 31 ~ "3",
              days > 31  & days <= 90 ~ "2",
              days > 90 ~ "1"))) |>
  select(state, sample_size, population, Republican, Democrat, weight)

IA_polls <- polls |>
  filter(state == "IA") |>
  filter(stage == "general") |>
  filter(answer %in% c("Turek", "Hinson")) |> 
  group_by(poll_id) |>
    filter(n_distinct(answer) > 1) |> 
    distinct(answer, .keep_all = TRUE) |> 
    slice_head(n = 2) |>
   ungroup() |>
pivot_wider(names_from = "answer", values_from = "pct") |>
  rename("Republican" = "Hinson") |>
  rename("Democrat" = "Turek") |>
  mutate(days = as.numeric(today() - date),
         weight = factor(
            case_when(
              days >= 0  & days <= 14 ~ "4",
              days > 14  & days <= 31 ~ "3",
              days > 31  & days <= 90 ~ "2",
              days > 90 ~ "1"))) |>
  select(state, sample_size, population, Republican, Democrat, weight)

KS_polls <- polls |>
  filter(state == "KS") |>
  filter(stage == "general") |>
  filter(answer %in% c("Marshall", "Hamilton")) |>
  group_by(poll_id) |>
    filter(n_distinct(answer) > 1) |> 
    distinct(answer, .keep_all = TRUE) |> 
    slice_head(n = 2) |>
   ungroup() |>
 pivot_wider(names_from = "answer", values_from = "pct") |>
  rename("Republican" = "Marshall") |>
  rename("Democrat" = "Hamilton") |>
  mutate(days = as.numeric(today() - date),
         weight = factor(
            case_when(
              days >= 0  & days <= 14 ~ "4",
              days > 14  & days <= 31 ~ "3",
              days > 31  & days <= 90 ~ "2",
              days > 90 ~ "1"))) |>
  select(state, sample_size, population, Republican, Democrat, weight)

ME_polls <- polls |>
  filter(state == "ME") |>
  filter(stage == "general") |>
  filter(answer %in% c("Jackson (Troy)", "Collins (Susan M.)")) |>
  group_by(poll_id) |>
    filter(n_distinct(answer) > 1) |> 
    distinct(answer, .keep_all = TRUE) |> 
    slice_head(n = 2) |>
   ungroup() |>
pivot_wider(names_from = "answer", values_from = "pct") |>
  rename("Democrat" = "Jackson (Troy)") |>
  rename("Republican" = "Collins (Susan M.)") |>
  mutate(days = as.numeric(today() - date),
         weight = factor(
            case_when(
              days >= 0  & days <= 14 ~ "4",
              days > 14  & days <= 31 ~ "3",
              days > 31  & days <= 90 ~ "2",
              days > 90 ~ "1"))) |>
  select(state, sample_size, population, Republican, Democrat, weight)

MI_polls <- polls |>
  filter(state == "MI") |>
  filter(stage == "general") |>
  filter(answer %in% c("El-Sayed", "Rogers (Mike)")) |>
  group_by(poll_id) |>
    filter(n_distinct(answer) > 1) |> 
    distinct(answer, .keep_all = TRUE) |> 
    slice_head(n = 2) |>
   ungroup() |>
pivot_wider(names_from = "answer", values_from = "pct") |>
  rename("Republican" = "Rogers (Mike)") |>
  rename("Democrat" = "El-Sayed") |>
  mutate(days = as.numeric(today() - date),
         weight = factor(
            case_when(
              days >= 0  & days <= 14 ~ "4",
              days > 14  & days <= 31 ~ "3",
              days > 31  & days <= 90 ~ "2",
              days > 90 ~ "1"))) |>
  select(state, sample_size, population, Republican, Democrat, weight)

MN_polls <- polls |>
  filter(state == "MN") |>
  filter(stage == "general") |>
  filter(answer %in% c("Flanagan (Peggy)", "Tafoya")) |>
  group_by(poll_id) |>
    filter(n_distinct(answer) > 1) |> 
    distinct(answer, .keep_all = TRUE) |> 
    slice_head(n = 2) |>
   ungroup() |>
pivot_wider(names_from = "answer", values_from = "pct") |>
  rename("Democrat" = "Flanagan (Peggy)") |>
  rename("Republican" = "Tafoya") |>
  mutate(days = as.numeric(today() - date),
         weight = factor(
            case_when(
              days >= 0  & days <= 14 ~ "4",
              days > 14  & days <= 31 ~ "3",
              days > 31  & days <= 90 ~ "2",
              days > 90 ~ "1"))) |>
  select(state, sample_size, population, Republican, Democrat, weight)

NC_polls <- polls |>
  filter(state == "NC") |>
  filter(stage == "general") |>
  filter(answer %in% c("Whatley", "Cooper")) |>
  group_by(poll_id) |>
    filter(n_distinct(answer) > 1) |> 
    distinct(answer, .keep_all = TRUE) |> 
    slice_head(n = 2) |>
   ungroup() |>
pivot_wider(names_from = "answer", values_from = "pct") |>
  rename("Democrat" = "Cooper") |>
  rename("Republican" = "Whatley") |>
  mutate(days = as.numeric(today() - date),
         weight = factor(
            case_when(
              days >= 0  & days <= 14 ~ "4",
              days > 14  & days <= 31 ~ "3",
              days > 31  & days <= 90 ~ "2",
              days > 90 ~ "1"))) |>
  select(state, sample_size, population, Republican, Democrat, weight)

NE_polls <- polls |>
  filter(state == "NE") |>
  filter(stage == "general") |>
  filter(answer %in% c("Osborn", "Ricketts")) |>
  group_by(poll_id) |>
    filter(n_distinct(answer) > 1) |> 
    distinct(answer, .keep_all = TRUE) |> 
    slice_head(n = 2) |>
   ungroup() |>
pivot_wider(names_from = "answer", values_from = "pct") |>
  rename("Democrat" = "Osborn") |>
  rename("Republican" = "Ricketts") |>
  mutate(days = as.numeric(today() - date),
         weight = factor(
            case_when(
              days >= 0  & days <= 14 ~ "4",
              days > 14  & days <= 31 ~ "3",
              days > 31  & days <= 90 ~ "2",
              days > 90 ~ "1"))) |>
  select(state, sample_size, population, Republican, Democrat, weight)

NH_polls <- polls |>
  filter(state == "NH") |>
  filter(stage == "general") |>
  filter(answer %in% c("Pappas", "Sununu")) |>
  group_by(poll_id) |>
    filter(n_distinct(answer) > 1) |> 
    distinct(answer, .keep_all = TRUE) |> 
    slice_head(n = 2) |>
   ungroup() |>
pivot_wider(names_from = "answer", values_from = "pct") |>
  rename("Democrat" = "Pappas") |>
  rename("Republican" = "Sununu") |>
  mutate(days = as.numeric(today() - date),
         weight = factor(
            case_when(
              days >= 0  & days <= 14 ~ "4",
              days > 14  & days <= 31 ~ "3",
              days > 31  & days <= 90 ~ "2",
              days > 90 ~ "1"))) |>
  select(state, sample_size, population, Republican, Democrat, weight)

OH_polls <- polls |>
  filter(state == "OH") |>
  filter(stage == "general") |>
  filter(answer %in% c("Brown", "Husted")) |>
  group_by(poll_id) |>
    filter(n_distinct(answer) > 1) |> 
    distinct(answer, .keep_all = TRUE) |> 
    slice_head(n = 2) |>
   ungroup() |>
pivot_wider(names_from = "answer", values_from = "pct")  |>
  rename("Democrat" = "Brown") |>
  rename("Republican" = "Husted") |>
  mutate(days = as.numeric(today() - date),
         weight = factor(
            case_when(
              days >= 0  & days <= 14 ~ "4",
              days > 14  & days <= 31 ~ "3",
              days > 31  & days <= 90 ~ "2",
              days > 90 ~ "1"))) |>
  select(state, sample_size, population, Republican, Democrat, weight)

TX_polls <- polls |>
  filter(state == "TX") |>
  filter(stage == "general") |>
  filter(answer %in% c("Talarico", "Paxton")) |>
  group_by(poll_id) |>
    filter(n_distinct(answer) > 1) |> 
    distinct(answer, .keep_all = TRUE) |> 
    slice_head(n = 2) |>
   ungroup() |>
pivot_wider(names_from = "answer", values_from = "pct") |>
  rename("Democrat" = "Talarico") |>
  rename("Republican" = "Paxton") |>
  mutate(days = as.numeric(today() - date),
         weight = factor(
            case_when(
              days >= 0  & days <= 14 ~ "4",
              days > 14  & days <= 31 ~ "3",
              days > 31  & days <= 90 ~ "2",
              days > 90 ~ "1"))) |>
  select(state, sample_size, population, Republican, Democrat, weight)

all_polls <- bind_rows(
  "AK" = AK_polls, 
  "FL" = FL_polls, 
  "GA" = GA_polls, 
  "IA" = IA_polls, 
  "KS" = KS_polls, 
  "ME" = ME_polls, 
  "MI" = MI_polls, 
  "MN" = MN_polls, 
  "NC" = NC_polls, 
  "NE" = NE_polls, 
  "NH" = NH_polls, 
  "OH" = OH_polls, 
  "TX" = TX_polls,
  .id = "state" # This creates a new column named "state" with your labels
)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
all_states_summary <- all_polls |>
  group_by(state) |> 
  mutate(
    R_num      = as.numeric(as.character(Republican)),
    D_num      = as.numeric(as.character(Democrat)),
    weight_num = as.numeric(as.character(weight)),
    
    R_W  = R_num * weight_num,
    D_W  = D_num * weight_num,
    ss_W = sample_size * weight_num
  ) |>
  summarize(
    Rs     = sum(R_W, na.rm = TRUE) / sum(weight_num, na.rm = TRUE),
    Ds     = sum(D_W, na.rm = TRUE) / sum(weight_num, na.rm = TRUE),
    ss    = sum(ss_W, na.rm = TRUE) / sum(weight_num, na.rm = TRUE),
    polls = n(),
    ME    = (.98 / sqrt(ss)) * 100,
    SD    = ME / 1.96,
    .groups = "drop" 
  ) |>
 mutate(
    total_pct = Rs + Ds,
    Rs_norm = (Rs / total_pct) * 100,
    Ds_norm = (Ds / total_pct) * 100
  )
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
All_Data <- CPVI |>
  right_join(all_states_summary) |>
  mutate(D_adj_pvi = 50 + adj_pvi / 2,
         R_adj_pvi = 50 - adj_pvi / 2,
         D_pvi = 50 + pvi / 2,
         R_pvi = 50 - pvi / 2) |>
  mutate(R = Rs_norm*(1-Nat_Env-Sta_Env) + R_adj_pvi*Nat_Env + R_pvi*Sta_Env,
         D = Ds_norm*(1-Nat_Env-Sta_Env) + D_adj_pvi*Nat_Env + D_pvi*Sta_Env)  |>
    mutate(D_margin = D-R,
           SD_margin = SD*2)
  

Some_Data <- All_Data |>
  select(state, R, D) |>
  rename("State" = "state")
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
run_state_sim <- function(state_df, iterations = 10000) {
  # Simulate margin directly using Normal Distribution
  sim_margin <- rnorm(iterations, mean = state_df$D_margin, sd = state_df$SD_margin)
  
  data.frame(
    State      = state_df$state,
    Republican = mean(sim_margin < 0),
    Democrat   = mean(sim_margin > 0)
  )
}

all_simulations <- do.call(rbind, lapply(split(All_Data, All_Data$state), run_state_sim, iterations = 10000)) |>
  mutate(D_win = ifelse(Democrat > Republican, 1, 0))
 
Dem_Wins <- sum(all_simulations$D_win)
rownames(all_simulations) <- NULL
```


```{r, echo = FALSE, message = FALSE, warning = FALSE}
states_sf <- read_sf("https://rstudio.github.io/leaflet/json/us-states.geojson")  #For interactive plot

states_int <- states_sf |>
  mutate(State = state.abb[match(name, state.name)]) |>
  mutate(State = if_else(name == "District of Columbia", "DC", State))

interactive <- states_int |>
  left_join(all_simulations) |>
  filter(!is.na(Democrat)) |>
  left_join(Some_Data) |>
  mutate(Dem_pct = if_else(Democrat <= 1, Democrat * 100, Democrat),
    Rep_pct = if_else(Republican <= 1, Republican * 100, Republican))

#Creating bins for coloring the plot
bins_party <- c(0, 20, 40, 48, 52, 60, 80, 100)
pal_party  <- colorBin("RdBu", domain = c(0, 100), bins = bins_party)
states <- interactive |>
  mutate(labels = str_c(State, ": ", Democrat, " Percent Chance of Winning"))
labels <- lapply(states$labels, HTML)

labels_party <- sprintf(
  "<div style='font-family: sans-serif;'>
     <strong style='font-size: 14px;'>%s (%s)</strong><br/>
     <span style='color: #2B5C8F;'>&#9632; Democrat Win Probability:</span> <b>%.1f%%</b><br/>
     <span style='color: #A61C1C;'>&#9632; Republican Win Probability:</span> <b>%.1f%%</b><br/>
     <hr style='margin: 5px 0;'/>
     <span style='color: #2B5C8F;'>&#9632; Democrat Vote Share:</span> <b>%.1f%%</b><br/>
     <span style='color: #A61C1C;'>&#9632; Republican Vote Share:</span> <b>%.1f%%</b>
   </div>",
  interactive$name,
  interactive$State,
  interactive$Dem_pct,
  interactive$Rep_pct,
  interactive$D, # Ensure columns match your dataset (D / R or Ds / Rs)
  interactive$R
) |> lapply(htmltools::HTML)


leaflet(interactive, width = "100%") |>
  setView(-84, 37.8, 4) |>
  addTiles() |>
  addPolygons(
    fillColor = ~pal_party(Dem_pct),
    weight = 2,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.8,
    highlightOptions = highlightOptions(
      weight = 2,
      color = "#333",
      dashArray = "",
      fillOpacity = 0.9,
      bringToFront = TRUE),
    label = labels_party,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "6px 10px"),
      textsize = "13px",
      direction = "auto")) |>
  addLegend(pal = pal_party, values = ~Dem_pct, opacity = 0.8, title = "Win Probability", position = "bottomright", labFormat = labelFormat(suffix = "%"))
```

Last Updated on September 16th, 2026

## Forecast Methodology: ##

These forecasts rely on polling data pulled from the New York Times’ database. This model aggregates each state’s polling data to find a race average, which places greater weights on recent polls to account for race momentum. 

The polls in this model account for 70% of the forecast’s weight, but that weight is easily modifiable in the forecast’s code. 

The remaining 30% of the forecast’s weight comes from a combination of the national political environment and each state's political environment. The national political environment currently gives Democrats a ~7% advantage. However, that advantage is different state-to-state. The Cook Political Voting Index estimates each state’s baseline political preference, providing a strong baseline expectation for how a state could vote in any given election. Accounting for a national advantage over each state's baseline provides an estimation for how well a Democrat and a Republican should do in any given state during the current election cycle. 

Using this 75/25 model, I ran 10,000 simulations of each race using the **Monte Carlo method**. I matched each simulation result between Democratic and Republican opponents against one another, crowning a victor in each simulation. Based on the total amounts of wins for both Democrats and Republicans, I estimated the probability of victory for each race, alongside an estimated vote share.


## With greater data availability, I would: ##

**1. Weigh polls differently, depending on the pollster.** In its current form, the forecast only assigns weights based on the recency of a poll, in order to account for momentum shifts in the race. If I had access to data regarding the reliability of each pollster, I would also assign stronger weights to polls conducted by trusted pollsters. 

**2. Integrate campaign “war chests” into the model.** Candidates with cash on hand use their money for advertisements to gain traction with voters. As election day approaches and campaigns dip into their savings, spent cash may translate into stronger poll numbers. Therefore, candidates with a stronger financial outlook than their opponents should expect to have stronger odds of winning a race.

**3. Account for historical polling errors.** Particularly in previous presidential elections, polls have been notably incorrect in 2016, 2020, and 2024, with fairly consistent errors in each of those years. The polling errors appear to be more significant in recent presidential years than in midterms, but if data clearly showed a trend in midterm years, I would account for that too. In the 2016, 2020, and 2024 elections, there were several significant deviations between polling averages and real vote share, including states relevant to this forecast:

**Florida:** Underestimated Republican vote share by 2.41% on average, overestimated Democratic by 2.38%.

**Georgia:** Underestimated Republican vote share by 0.38% on average, overestimated Democratic by 0.35%

**Iowa:** Underestimated Republican vote share by 2.66% on average, overestimated Democratic by 3.46%

**Michigan:** Underestimated Republican vote share by 1.85% on average, overestimated Democratic by 1.96%

**Minnesota:** Underestimated Republican vote share by 0.66% on average, overestimated Democratic by 2.21%

**New Hampshire:** Underestimated Republican vote share by 1.41% on average, overestimated Democratic by 1.49%

**North Carolina:** Underestimated Republican vote share by 1.17% on average, overestimated Democratic by 1.83%

**Ohio:** Underestimated Republican vote share by 2.73% on average, overestimated Democratic by 2.62%

**Texas:** Underestimated Republican vote share by 1.78% on average, overestimated Democratic by 1.77%



[Data source for polls](https://www.nytimes.com/interactive/polls/congressional-vote-2026.html)

[Data source for Political Index](https://en.wikipedia.org/wiki/Cook_Partisan_Voting_Index)